### 2019-08-17 reports/article.tex
abstract in landscape ecology format

\begin{abstract}
  Urbanization is currently a global phenomenon that has become the most important form of landscape change and is increasingly affecting biodiversity and ecosystem functions.
In order to evaluate the impacts of urbanization and inform urban planning, it is important to understand the spatiotemporal patterns of land use change associated to urbanization.
This paper exploits three different frameworks, namely landscape metrics, urban growth modes and fractal analysis to characterize the spatiotemporal patterns of urbanization of the Swiss urban agglomerations of Zurich, Bern and Lausanne.
The land use inventory provided by the Swiss Federal Statistical Office was used to assemble four temporal snapshots from 1980 to 2016 at the extent of the urban agglomerations.
The time series of landscape metrics generally supports the diffusion and coalescence model of urban growth, with Zurich exhibiting most characteristics of coalescence while Bern and Lausanne seem to be at the transition between diffusion and coalescence.
Nevertheless, the analysis of the urban growth modes suggest that leapfrog development occurs at all periods, which contributes to an increasing fragmentation of natural patches and maintains the fractal configuration of the landscape.
The discussion reviews potential explanations for the observed landscape changes, and concludes with some planning implications.

>>>>

\begin{abstract}
  \hfill\break

  \emph{Context} % Urbanization is a global phenomenon that has become the most important form of landscape change and is increasingly affecting biodiversity and ecosystem functions.
  Urbanization is the most important form of landscape change and is increasingly affecting biodiversity and ecosystem functions.
  Understanding how landscape patterns change in space and time is central to the evaluation of the environmental impacts of urbanization. % and inform urban planning.

  \emph{Objectives} This research explores the spatiotemporal patterns of land use change in the urban agglomerations of Zurich, Bern and Lausanne, and compares them to prominent hypotheses of urbanization patterns and to concepts from the complexity sciences. The discussion reviews potential explanations for the observed landscape changes and planning implications.

  \emph{Methods} For each urban agglomeration, four temporal snapshots from 1980 to 2016 have been derived from the land use inventory of the Swiss Federal Statistical Office. A combination of landscape metrics, urban growth modes and fractal analysis serves to characterize the spatiotemporal patterns of urbanization.

  \emph{Results} The time series of landscape metrics generally supports the diffusion and coalescence model of urban growth, with Zurich exhibiting characteristics of coalescence while Bern and Lausanne seem to be at the transition between diffusion and coalescence.
  However, the urban growth modes reveal that leapfrog development occurs at all periods, increasingly fragmenting natural patches while maintaining the fractal configuration of the landscape.
  % The way in which the structural complexity of the landscapes is sustained throughout urbanization is seemingly analogous to dynamic behaviors encountered in a wide range of complex self-organizing systems.

  \emph{Conclusions} The combination of landscape metrics, growth modes, and fractal analysis provides novel insights that are hardly noticeable with any of the approaches individually. The observed landscape complexity is likely related to the distinct planning policies adopted by the various municipalities that configure each urban agglomeration, as well as to residential preferences for low density environments.
  % Current Federal planning precepts aiming at the concentration growth within the existing urban agglomerations should not overlook the valuable ecosystem services that patches of natural land can provide to urban dwellers.
